---
title: "Geospatial Analysis"
author: "Helene Angot"
date: "2/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r message=FALSE}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(leaflet)
library(scales)
```

```{r}
ak_shp_sf = read_sf("shapefiles/ak_regions_simp.shp")
```

```{r}
plot(ak_shp_sf)
```

To get the coordinate reference system:

```{r}
st_crs(ak_shp_sf)
```

To change the projection system used the following.
EPSG codes [here](http://spatialreference.org/ref/epsg/?page=1).

```{r}
ak_shp_sf = ak_shp_sf %>% 
  st_transform(crs = 3338)

st_crs(ak_shp_sf)
```

```{r}
plot(ak_shp_sf)
```

# Spatial Joins

```{r}
pop = read.csv("shapefiles/alaska_population.csv")
```

We need to transform .csv to sf file.
When we transform it, we give the original crs code.
Later we change to something else (i.e., same as above to be able to join them).

```{r}
pop_sf = st_as_sf(pop,
                  coords = c("lng", "lat"),
                  crs = 4326,
                  remove = FALSE) %>% 
  st_transform(crs = 3338)
```

```{r}
pop_joined_sf = st_join(pop_sf, ak_shp_sf, join = st_within)

plot(pop_joined_sf["region"])
```

# Group by and summarize

```{r}
pop_region = pop_joined_sf %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
```

```{r}
ak_pop_sf = left_join(ak_shp_sf, pop_region)

plot(ak_pop_sf["total_pop"])
```

```{r}
ak_mgmt = ak_pop_sf %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop))

plot(ak_mgmt["total_pop"])
```

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp")
```

# Static plots

```{r}
ggplot(ak_pop_sf) +
  geom_sf(data = ak_pop_sf, aes(fill = total_pop)) +
  geom_sf(data = pop_sf, size = 0.5) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)
```

# Interactive maps with leaflet

Section 19.6 of the [tutorial](http://training.arcticdata.io/materials/arctic-data-center-training/spatial-vector-analysis-using-sf.html).

In leaflet, we must use **unprojected** data.

```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

```{r}
ak_crs = st_transform(ak_pop_sf, crs = 4326)
pop_crs = st_transform(ak_pop_sf, crs = 4326)
```

```{r}
pal = colorNumeric(palette = "Reds", domain = ak_crs$total_pop)

m = leaflet(options = leafletOptions(crs = epsg3338)) %>% 
  addPolygons(data = ak_crs,
              fillColor = ~pal(total_pop),
              weight = 1,
              color = "black",
              label = ~region,
              fillOpacity = 1) 
#%>% 
 # addCircleMarkers(data = pop_crs,
  #                 lat = ~lat,
   #                lng = ~lng,
    #               radius = 2,
     #              fillColor = "gray",
      #             fillOpacity = 1,
       #            weight = 0.25,
        #           color = "black",
         #          label = ~city)

m
```