---
title: "Data visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. Data [here](https://knb.ecoinformatics.org/view/doi:10.5063/F1S46Q6M).

# Packages

```{r message=FALSE, warning=FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```

# Reformat the data for plotting

```{r}
esc = read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = FALSE)
```

* calculate annual escapment by species and region

First we need to create a "Year" variable.

```{r}
dates_split = esc %>% 
  separate(sampleDate, into = c("year","month","day"), sep = "-", remove = F)
head(dates_split)

# coerce year to numeric type
dates_split = dates_split %>% 
  mutate(year = as.numeric(year))
head(dates_split)
```

Then we use functions "group_by", "summarize", and "sum" to calculate the annual escapment by species and region.

```{r}
annual_esc = dates_split %>% 
  group_by(year, Species, SASAP.Region) %>% 
  summarize(count = sum(DailyCount))

head(annual_esc)
```

* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

To do that we use function "filter".

```{r}
annual_esc = annual_esc %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Coho", "Pink", "Chum")) 
head(annual_esc)
```

# Generate some static plots

```{r}
ggplot(annual_esc, mapping = aes(x = Species, y = count, fill = SASAP.Region)) +
  geom_col()
```

```{r}
my_theme = theme_test() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement (num of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```

# Generate an interactive map

